---
title: "Assemblage quality control report"
date: "`r Sys.Date()`"
output:
  html_document:
    mathjax: null
    code_folding: "hide"
    toc: true
    toc_float:
      collapsed: false
params:
  quast_report:
    value: x
  kraken_report:
    value: x
  coverage_report:
    value: x
  genome_size_val:
    value: x
  species_name:
    value: x
---

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(kableExtra)
library(ggpubr)
library(formattable)
library(sparkline)
```


# QUAST assembly quality check

QUAST is used to run general quality control on assemblies by investigating total size, N50, %GC, number of contigs, and other parameters. 
```{r, message=FALSE, warning=FALSE, dpi=600, fig.width=10, fig.height=6, fig.cap="Figure 1: QUAST assembly general statistics."}
quast_data <- read_delim(
  params$quast_report,
  delim = "\t"
  ) %>%
  select(Assembly, `# contigs`, `Total length`, `GC (%)`, N50)

# Generate data frame in long format from the data, used for plotting
plot_data <- quast_data %>%
  # Rename columns with 'rename' from dplyr
  rename("Number of contigs" = `# contigs`,
         "Total length of assembly" = `Total length`,
         "Percent GC" = `GC (%)`,
         "Assembly N50" = N50) %>%
  # Gather all columns except Assembly, and generate long format
  # with 'pivot_longer' from tidyr
  pivot_longer(cols = -Assembly,
               names_to = "key",
               values_to = "value")

# Specify a color palette for the plotting, one HEX color code for
# each variable plotted
palette <- c("Assembly N50" = "#fdbf6f",
             "Total length of assembly" = "#a6cee3",
             "Percent GC" = "#fb9a99",
             "Number of contigs" = "#b2df8a")

# Adjust options to plot full numbers in the plot, not scientific 
# notation (i.e. 10e+10 etc.)
options(scipen = 999)

# Plot data with 'ggplot' from ggplot2
# Specify x and y axes data inside 'aes'
p <- ggplot(plot_data, aes(key, value, fill = key)) +
  # Plot a violin plot with given data
  geom_violin(trim = FALSE) +
  # Plot a boxplot over the violin plot
  geom_boxplot(width = 0.05,
               fill = "white") +
  # Specify a minimal theme of the plot
  theme_minimal() +
  # Adjust the theme with specific options
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        # Remove legend
        legend.position = "none") +
  # Specify the color palette used
  scale_fill_manual(values = palette) +
  # Create sub-plots within the plot with their own axes
  facet_wrap(~key, nrow = 2, scales = "free")

axis_text_size <- 6
axis_title_size <- 8
dot_size <- 1

# Generate dotplots
p1 <- quast_data %>%
  ggplot(aes(N50, `Total length`)) +
  geom_point(size = dot_size) +
  labs(x = "Assembly N50",
       y = "Total length of assembly") +
  theme_bw() +
  theme(axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_text_size))

p2 <- quast_data %>%
  ggplot(aes(N50, `# contigs`)) +
  geom_point(size = dot_size) +
  labs(x = "Assembly N50",
       y = "Number of contigs") +
  theme_bw() +
  theme(axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_text_size))

p3 <- quast_data %>%
  ggplot(aes(`# contigs`, `Total length`)) +
  geom_point(size = dot_size) +
  labs(x = "Number of contigs",
       y = "Total length of assembly") +
  theme_bw() +
  theme(axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_text_size))

p4 <- quast_data %>%
  ggplot(aes(`GC (%)`, `Total length`)) +
  geom_point(size = dot_size) +
  labs(x = "Percent GC",
       y = "Total length of assembly") +
  theme_bw() +
  theme(axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_text_size))

# Combine and arrange the plots with 'ggarrange' from ggpubr
p_a <- ggarrange(p1, p2, p3, p4, nrow = 2, ncol = 2)

# Finally, combine both plots into one
ggarrange(p, p_a, nrow = 1, labels = c("A","B"))
```

## Genome size check

```{r}
genome_size <- as.numeric(params$genome_size_val)

test <- quast_data %>%
  mutate(test = ifelse(
    `Total length` > genome_size * 1.2 | `Total length` < genome_size * 0.8,
    FALSE,
    TRUE
  )) %>%
  filter(test == FALSE)

if (nrow(test) == 0) {
  "All genomes are within set genome size limits."
} else {
  test %>%
    select(-test) %>%
    kbl(escape = FALSE, caption = "List of genomes outside of genome size limits.") %>%
    # Specify the look of the table
    kable_paper("hover", full_width = TRUE)
}


  
```


# Kraken contamination check
Kraken2 is a tool that maps reads to a database of a myriad of bacterial species, and gives the percent of reads that mapped to each taxonomic level. The tool is useful when looking for potential contamination in sequenced samples.

```{r, message=FALSE, warning=FALSE, dpi=600, fig.width=10, fig.height=5, fig.cap="Figure 2: Top species hits in the Kraken database (>0.1% of fragments)"}

kraken_data <- read_delim(
  params$kraken_report,
  delim = "\t",
  col_names = TRUE
) %>%
  mutate(perc_fragments = as.numeric(perc_fragments)) %>%
  filter(rank_code == "S",
         perc_fragments > 0.1) %>%
  select(ref, scientific_name, perc_fragments) %>%
  rename("Sample" = ref)

ggplot(kraken_data, aes(reorder(scientific_name, -perc_fragments), perc_fragments)) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot(fill = "grey80") +
  labs(x = "Species",
       y = "Percent fragments") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3))

```

<br>

<br>

# Coverage calculation
Coverage calculation is used to calculate how many reads are each base in the assembly is covered by. Assemblies with a very low coverage depth means that each nucleotide has low confidence, and the assembly should be treated with caution.

```{r, message=FALSE, warning=FALSE}
sparkline(0)

read_delim(
  params$coverage_report,
  delim = "\t",
  col_names = TRUE
  ) %>%
  kbl(escape = FALSE, caption = "Actual coverage statistics for all assemblies.") %>%
  # Specify the look of the table
  kable_paper("hover", full_width = TRUE) %>%
  # Adjust the first column text
  column_spec(1, bold = TRUE, color = "grey") %>%
  column_spec(5, width = "4cm") %>%
  # Add a scroll box
  scroll_box(height = "400px")

```
